---
title: "MeDNA Survey123 Summary"
output: html_notebook
---

# STEP 1 - Read in CSVs from Survey123

This notebook:
1. Loads Survey123 data
2. Joins tables based on primary and foreign keys
3. Generates "long" tables by melting repeats


```{r, echo=FALSE, results="hide", messages=FALSE}
# Load and Install Libraries
source("medna_survey123.R")
## Check libraries & install
LibraryList<-c("data.table","dplyr","lubridate","stringr","ggplot2","scales", "zoo", "tidyr")
install_or_load_pack(LibraryList)
overwrite = TRUE
inputFolder = "../data/02_Working/PyOutput/"
outputFolder = "../data/02_Working/ROutput/"

outputPngFolder = paste0(outputFolder,"Figures/")

todaysDateFn = format(Sys.Date(), "%Y%m%d")
```

# load in data
```{r}
#load data
survey_sub <- data.table::fread(paste0(inputFolder,'eDNA_Sampling_v14_sub.csv'))
survey_envmeas_join <- data.table::fread(paste0(inputFolder,'survey_envmeas_join.csv'))
survey_collection_join <- data.table::fread(paste0(inputFolder,'survey_collection_join.csv'))
clean_filter_join <- data.table::fread(paste0(inputFolder,'clean_filter_join.csv'))
```
# review data
```{r}
# displays column names & number of rows - 6589
names(survey_sub);nrow(survey_sub)
names(survey_envmeas_join);nrow(survey_envmeas_join)
names(survey_collection_join);nrow(survey_collection_join)
names(clean_filter_join);nrow(clean_filter_join)
# Head displays the first 6 rows of the data.table
#head(survey_sub);head(survey_envmeas_join);head(survey_collection_join);head(clean_filter_join);head(clean_filter_join)
```

# check empty site_ids
```{r}
# How do you want to treat NA?
# subset the data by the NAs, explore them & their origin
# can we programmatically fix it, or do we have to do it manually?

# 0 rows blank/NA Site IDs
survey_empty<-survey_sub[survey_sub$site_id == "" | is.na(survey_sub$site_id),]
nrow(survey_empty)
rm(survey_empty)
```


"survey_GlobalID"       "survey_date"           "survey_month"         
"survey_year"           "projects"              "supervisor"            "username"             
"recorder_first_name"   "recorder_last_name"    "system_type"           "site_id"              
"other_site_id"         "general_location_name" "envmeas_date"          "envmeas_depth"        
"envmeas_instrument"    "ctd_filename"          "ctd_notes"             "ysi_filename"         
"ysi_model"             "ysi_serial_number"     "ysi_notes"             "secchi_depth"         
"secchi_notes"          "niskin_number"         "niskin_notes"          "other_instruments"    
"env_measurements"      "flow_rate"             "water_temp"            "salinity"             
"ph"                    "par1"                  "par2"                  "turbidity"            
"conductivity"          "do"                    "pheophytin"            "chla"                 
"no3no2"                "no2"                   "nh4"                   "phosphate"            
"bottom_substrate"      "lab_date"              "envmeas_notes"         "envmeas_GlobalID"     
"survey_edit_date"      "survey_create_date"    "envmeas_create_date"  

```{r}

unique_survey_gids <- unique(survey_sub$survey_GlobalID)
unique_envmeas_survey_gids <- unique(survey_envmeas_join$survey_GlobalID) 

perc_gids<-(length(unique_envmeas_survey_gids)/length(unique_survey_gids))*100
print(paste0("Percentage of survey records with environmental measurements: ",round(perc_gids,2),"%"))

unique_survey_sids <- unique(survey_sub$site_id)
unique_envmeas_sids <- unique(survey_envmeas_join$site_id)

perc_sids<-(length(unique_envmeas_sids)/length(unique_survey_sids))*100
print(paste0("Percentage of site ids with environmental measurements: ",round(perc_sids,2),"%"))


nrow(survey_envmeas_join)
survey_envmeas_join %>%
         dplyr::select(envmeas_GlobalID, site_id) %>% 
         dplyr::group_by(site_id) %>%
         dplyr::count(site_id) %>%
         dplyr::mutate(pct=n/sum(n))

ggplot(survey_envmeas_join %>%
         dplyr::select(envmeas_GlobalID, site_id) %>% 
         dplyr::group_by(site_id) %>%
         dplyr::count(envmeas_GlobalID) %>%
         dplyr::mutate(pct=n/sum(n)),              # Calculate percent within each region
       aes(site_id, n, fill=site_id)) +
  geom_bar(stat="identity") +
  xlab("type") +
  ggtitle("Count of Survey123 records in freezer by type") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  theme(legend.position = "none") +
  guides(fill=guide_legend(title="In Freezer"))
```


# unique survey records & missing env measurements
```{r}
fileName="s123_in_freezer_count_pff"
outputPngFileName <- paste0(outputPngFolder,fileName,".png")
png(outputPngFileName,height=4,width=6, units='in', res=300)
ggplot(survey123_all %>%
         dplyr::select(in_freezer,filter_prefilter) %>% 
         dplyr::count(in_freezer,filter_prefilter) %>%
         dplyr::group_by(filter_prefilter) %>%
         dplyr::mutate(filter_prefilter,pct=n/sum(n)) %>%     
         dplyr::arrange(in_freezer),              # Calculate percent within each region
       aes(filter_prefilter, n, fill=in_freezer)) +
  geom_bar(stat="identity") +
  xlab("type") +
  ggtitle("Count of Survey123 records in freezer by type") +
  theme(plot.title = element_text(hjust = 0.5)) +
  guides(fill=guide_legend(title="In Freezer")) +
  scale_fill_manual("In Survey123", values = c("?" = "#CC79A7", "Y" = "#009E73", "N" = "#FF6666")) +
  geom_text(aes(label=paste0(sprintf("%1.1f", pct*100),"%"," (",n,")")), 
            position=position_stack(vjust=0.5))
print(paste("Plot saved as:",outputPngFileName))
dev.off()


```

```{r}
# write full join (no clean) to file
outputFileName = "preclean_prefilter_filter_join"
outputFile = paste0(outputFolder,outputFileName,".csv")

if (!file.exists(outputFile) | overwrite) {
  write.csv(preclean_prefilter_filter_join, outputFile, row.names=FALSE)
}
```
